---
title: "Compile data for Rio de Janeiro Municipality between 2012-2016"
author: "Marcelo Ferreira da Costa Gomes"
date: "3 de fevereiro de 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading data

We'll use dengue data from Rio de Janeiro Municipality from 2012 to 2016, taken from the DENGON database. Since the 2016 dataset have columns that differ from those of past years, we will read only the relevant and consistent ones.
```{r, warning=FALSE, message=FALSE}
library(foreign)
target.cols <- c('NU_NOTIFIC', 'ID_MUNICIP', 'ID_UNIDADE', 'DT_NOTIFIC', 'DT_DIGITA')
flist <- list.files('./data/', pattern='*.dbf$', full.names = T)
df.dbf <- read.dbf(flist[1], as.is=T)[, target.cols]
for (fname in flist[2:length(flist)]){
  df.dbf <- rbind(df.dbf, read.dbf(fname, as.is=T)[, target.cols])
}
geocod <- 330455
df.dbf <- df.dbf[df.dbf$ID_MUNICIP == geocod, ]
```

Filter by list of columns of interest and drop possible duplicates:
```{r}
filter.cols <- c('NU_NOTIFIC', 'ID_MUNICIP', 'ID_UNIDADE', 'DT_NOTIFIC')
df.dbf.clean <- df.dbf[(!duplicated(df.dbf[, filter.cols]) &
                                !is.na(df.dbf$DT_DIGITA)), ]

nrow(df.dbf.clean)
```

Next we are going to generate epidemiological week information to both notification and digitization dates
```{r, message=FALSE}
require(leos.opportunity.estimator)
# Create columns with epiweek, epiyear, and epiyearweek from notification and digitalization ones:
target.cols <- c('DT_NOTIFIC_epiyearweek', 'DT_NOTIFIC_epiweek', 'DT_NOTIFIC_epiyear')
if (!all(target.cols %in% names(df.dbf.clean))){
  df.dbf.clean <- generate.columns.from.date(df.dbf.clean, 'DT_NOTIFIC')
  names(df.dbf.clean) <- sub("^epi", "DT_NOTIFIC_epi", names(df.dbf.clean))
}
target.cols <- c('DT_DIGITA_epiyearweek', 'DT_DIGITA_epiweek', 'DT_DIGITA_epiyear')
if (!all(target.cols %in% names(df.dbf.clean))){
  df.dbf.clean <- generate.columns.from.date(df.dbf.clean, 'DT_DIGITA')
  names(df.dbf.clean) <- sub("^epi", "DT_DIGITA_epi", names(df.dbf.clean))
}
names(df.dbf.clean)
```

Save data on disk:
```{r}
saveRDS(df.dbf.clean, file='data/dengue.munRJ.dbf.2012.2016.epiyearweek.rds')
write.csv(df.dbf.clean, file='data/dengue.munRJ.dbf.2012.2016.epiyearweek.csv', row.names=F)
```